The Conjunctive Disjunctive Graph Node Kernel for Disease Gene Prioritization

### 2017-11-16 Response_neurocomputing.tex

The proposed kernel takes labeled graphs in input. The appropriate node labeling function is domain dependent and can be freely defined by the user. In this paper we propose a couple of labeling functions to evaluate the performance of the disease gene prioritization system with the use of our kernel. We consider a trivial constant label to show the capacity of the approach to exploit topological information only. In addition we consider a domain dependent source of information derived by the GO hierarchy. We agree that a more extensive evaluation for labeling functions would improve the performance on a specific task. The scope of the evaluation presented in this paper is however only to show that CDNK can in principle improve results by exploiting domain knowledge.

\subsection*{3} \textit{The paper is extremely badly organized, with some notions (like conjunctive and disjunctive edges) used before they are defined. I suggest the following re-organization of the paper:}

\begin{enumerate}
\item \textit{Introduction (that must include a discussion about the problem at stake, gene prioritization)}

\underline{Answer}: we have added the discussion in the new version of the paper.

@@

\item \textit{Related to the previous comment, the notion of conjunctive and disjuctive edges seems central, however they are shortly mentioned in Section 2.2.2 and then shorlty re-defined (?) in Section 2.4. In the first case, the description of these edges must be expanced and clarified, togheter with some more details about Figure 1 (caption and text). Maybe it can be anticipated in the "Definitions" paragraph (page 3), or in the "initial decription" that I introduced in the previous comment.}

\underline{Answer}: We have added a more detailed explanation in the new version of the paper based on the reviewer's comment.

\item \textit{Overall, there is very large number of parameters involved in the proposed approach. The authors describe their model validation procedure in Page 9/10, however the space of possible combinations of parameters is huge + the choices over other thresholds described in the paper. This seems to be an important drawback of the method, and some comments on this point should be added.}

\underline{Answer}: We have added the following section named ``Hyper-parameter" to explain why the parameter space for the proposed kernel is not high.\\

### 2017-11-16 neurocomputingv3.tex
F

\subsection{Additional Node Information}
In order to integrate the information of real vector labels that may be present associated to nodes, we proceed as follows. We compute a sparse vector representation for the conjunctive neighborhood graph rooted at node $u$ following \cite{costa2010fast}: for each conjunctive neighborhood subgraph we calculate the quasi-isomorphism certificate hash code; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator. This yields a direct sparse vector representation (associated to node $u$ in graph $G^{\wedge\vee}$) $f: u \longmapsto \mathbb{R}^N$ where $N \approx 10^6$. Given the real valued vector information associated to nodes in graph $G^{\wedge\vee}$, we also cluster\footnote{Any clustering algorithm that returns cluster representations provided of a function to measure the similarity of a vector with respect to a given cluster will do the job.} them in $P$ clusters (c.f.r. Section \ref{sec:lab}) and define a function $g: u \longmapsto \mathbb{R}^P$ which returns the multi-class similarity w.r.t. each of the $P$ cluster (prototypes).

@@

\textit{Real vector labels:} In addition to encoding the functional information as a discrete label, we add a richer description by computing the similarity vector w.r.t. to each cluster. In this way, we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure. Formally, given a vector $v \in \mathbb{R}^{M}$ \footnote{M is the number of GO-terms, M = 26,501} we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.

\subsection{Model Selection}
The hyper-parameters of the various methods were tuned using a 10-fold cross validation. However, due to the non i.i.d. nature of the problem, we emploied a stronger setup to ensure no information leakage. The dataset $0$ (Connective) on which we  validated the performance was never subsequently used in the predictive performance estimation. The values for the diffusion parameter $\beta$ in LEDK and MEDK were sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$. For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 1, 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$. Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$. The optimal parameter values of CDNK obtained by validation process are $D = 1$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
%\subsection{Parameter Impact}
%In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 on the BioGPS network. For each parameter, first we fix the values of all other parameters, but allow the value of considered papameter to be changed. More precisely, we use the optimal tuple $C = 5,\ d^* = 1,\ D = 20,\ P = 15$ for the fixed of parameters' values. We let the changes of $C$ in $\lbrace 4,\ 5,\ 6,\ 7,\ 8,\ 9\rbrace$, $d^*$ in $\lbrace 1,\ 2,\ 3,\ 4\rbrace$, $D$ in $\lbrace 8,\ 10,\  15,\ 20,\ 30,\ 40,\ 50\rbrace$ and real vector size, $P$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 50,\ 70, 90\rbrace$.

@@

\section{Conclusion and Future Work}
We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results for a disease gene prioritization task.
Moreover, we have also introduced the possibility to integrate in the kernel computation  additional information in the form of real valued vectors in the case in which it is available for graph nodes. We experimentally showed that this possibility actually leads to even better performances. Future work we will investigate on how to: \textit{i}) decompose networks in a data driven way; \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.

@@

\bibliography{mybibfile}
\end{document}

>>>>

\subsection{Additional Node Information}
In order to integrate the information of real vector labels that may be present associated to nodes, we proceed as follows. We compute a sparse vector representation for the conjunctive neighborhood graph rooted at node $u$ following \cite{costa2010fast}: for each conjunctive neighborhood subgraph we calculate a quasi-isomorphism certificate hash code\footnote{The graph hashing is obtained by  hashing the sorted sequence of edge extended information which includes an invariant encoding of the endpoint labels (see \cite{costa2010fast}).}; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator. This yields a direct sparse vector representation (associated to node $u$ in graph $G^{\wedge\vee}$) $f: u \longmapsto \mathbb{R}^N$ where $N \approx 10^6$. Given the real valued vector information associated to nodes in graph $G^{\wedge\vee}$, we also cluster\footnote{Any clustering algorithm that returns cluster representations provided of a function to measure the similarity of a vector with respect to a given cluster will do the job.} them in $P$ clusters (c.f.r. Section \ref{sec:lab}) and define a function $g: u \longmapsto \mathbb{R}^P$ which returns the multi-class similarity w.r.t. each of the $P$ cluster (prototypes).

@@

\textit{Real vector labels:} In addition to encoding the functional information as a discrete label, we add a richer description by computing the similarity vector w.r.t. to each cluster. In this way, we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure. Formally, given a vector $v \in \mathbb{R}^{M}$ \footnote{M is the number of GO-terms, in our experiments M = 26,501.} we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$. While other notions of similarity can be used (e.g. the cosine similarity or a similarity derived from a more flexible  Mahalanobis distance), here we choose the Euclidean distance given the prior knowledge on the domain: the cosine similarity is often used as an effective way to normalize bag of words representations that encode frequencies or counts (while here we have a 0-1 encoding), and Mahalanobis distances are useful when unaccounted correlations and collinearities between features are expected (while in our case the ontology terms form a controlled non redundant vocabulary).

\subsection{Model Selection}
The hyper-parameters of the various methods were tuned using a 10-fold cross validation. However, due to the non i.i.d. nature of the problem, we emploied a stronger setup to ensure no information leakage. The dataset $0$ (Connective) on which we  validated the performance was never subsequently used in the predictive performance estimation. The values for the diffusion parameter $\beta$ in LEDK and MEDK were sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$. For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 1, 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$. Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$. The optimal parameter values of CDNK obtained by validation process are $D = 10$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
%\subsection{Parameter Impact}
%In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 on the BioGPS network. For each parameter, first we fix the values of all other parameters, but allow the value of considered papameter to be changed. More precisely, we use the optimal tuple $C = 5,\ d^* = 1,\ D = 20,\ P = 15$ for the fixed of parameters' values. We let the changes of $C$ in $\lbrace 4,\ 5,\ 6,\ 7,\ 8,\ 9\rbrace$, $d^*$ in $\lbrace 1,\ 2,\ 3,\ 4\rbrace$, $D$ in $\lbrace 8,\ 10,\  15,\ 20,\ 30,\ 40,\ 50\rbrace$ and real vector size, $P$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 50,\ 70, 90\rbrace$.

@@

\section{Conclusion and Future Work}
We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results for a disease gene prioritization task.
Moreover, we have also introduced the possibility to integrate in the kernel computation  additional information in the form of real valued vectors when such information is available for each graph node. We experimentally showed that this actually leads to improved performances. In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources.

@@

%\bibliography{mybibfile}
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\expandafter\ifx\csname href\endcsname\relax
  \def\href#1#2{#2} \def\path#1{#1}\fi

\bibitem{mordelet2011prodige}
F.~Mordelet, J.-P. Vert, Prodige: Prioritization of disease genes with
  multitask machine learning from positive and unlabeled examples, BMC
  bioinformatics 12~(1) (2011) 389.

\bibitem{chen2014disease}
B.~Chen, M.~Li, J.~Wang, F.-X. Wu, Disease gene identification by using graph
  kernels and markov random fields, Science China. Life Sciences 57~(11) (2014)
  1054.

\bibitem{valentini2014extensive}
G.~Valentini, A.~Paccanaro, H.~Caniza, A.~E. Romero, M.~Re, An extensive
  analysis of disease-gene associations using network integration and fast
  kernel-based gene prioritization methods, Artificial Intelligence in Medicine
  61~(2) (2014) 63--78.

\bibitem{kondor2002diffusion}
R.~I. Kondor, J.~Lafferty, Diffusion kernels on graphs and other discrete input
  spaces, in: ICML, Vol.~2, 2002, pp. 315--322.

\bibitem{fouss2006experimental}
F.~Fouss, L.~Yen, A.~Pirotte, M.~Saerens, An experimental investigation of
  graph kernels on a collaborative recommendation task, in: Data Mining, 2006.
  ICDM'06. Sixth International Conference on, IEEE, 2006, pp. 863--868.

\bibitem{chebotarev2006matrix}
P.~Chebotarev, E.~Shamis, The matrix-forest theorem and measuring relations in
  small social groups, arXiv preprint math/0602070.

\bibitem{haussler1999convolution}
D.~Haussler, Convolution kernels on discrete structures, Tech. rep., Technical
  report, Department of Computer Science, University of California at Santa
  Cruz (1999).

\bibitem{gartner2003survey}
T.~G{\"a}rtner, A survey of kernels for structured data, ACM SIGKDD
  Explorations Newsletter 5~(1) (2003) 49--58.

\bibitem{borgwardt2005shortest}
K.~M. Borgwardt, H.-P. Kriegel, Shortest-path kernels on graphs, in: Data
  Mining, Fifth IEEE International Conference on, IEEE, 2005, pp. 8--pp.

\bibitem{shervashidze2009fast}
N.~Shervashidze, K.~M. Borgwardt, Fast subtree kernels on graphs, in: Advances
  in neural information processing systems, 2009, pp. 1660--1668.

\bibitem{shervashidze2011weisfeiler}
N.~Shervashidze, P.~Schweitzer, E.~J.~v. Leeuwen, K.~Mehlhorn, K.~M. Borgwardt,
  Weisfeiler-lehman graph kernels, Journal of Machine Learning Research
  12~(Sep) (2011) 2539--2561.

\bibitem{costa2010fast}
F.~Costa, K.~De~Grave, Fast neighborhood subgraph pairwise distance kernel, in:
  Proceedings of the 26th International Conference on Machine Learning,
  Omnipress, 2010, pp. 255--262.

\bibitem{alvarez2005k}
J.~I. Alvarez-Hamelin, L.~Dall'Asta, A.~Barrat, A.~Vespignani, k-core
  decomposition: A tool for the visualization of large scale networks, arXiv
  preprint cs/0504107.

\bibitem{tarjan1985decomposition}
R.~E. Tarjan, Decomposition by clique separators, Discrete mathematics 55~(2)
  (1985) 221--232.

\bibitem{goh2007human}
K.-I. Goh, M.~E. Cusick, D.~Valle, B.~Childs, M.~Vidal, A.-L. Barab{\'a}si, The
  human disease network, Proceedings of the National Academy of Sciences
  104~(21) (2007) 8685--8690.

\bibitem{gene2004gene}
G.~O. Consortium, et~al., The gene ontology (go) database and informatics
  resource, Nucleic acids research 32~(suppl 1) (2004) D258--D261.

\end{thebibliography}

\end{document}
